---
title: "UBC Science Specialization Data"
output: html_document
---

## Personal Project - UBC Science Specializations Analysis

## Author
* Calvin Du (calvingdu@gmail.com)

# Introduction
In this project, I'll be looking at the UBC Science Specializations Spreadsheet as provided by a FOI request to UBC. I'll be exploring the data in various ways in order to determine if COVID was a factor in that in the increasing grade requirements for specializations in 2020 & 2021. 

# Part 1: Loading and Cleaning up Data 

To begin, I first load the data and clean it up to allow for consistency and to remove any errors. 

```{r}
library(tidyverse)
library(haven)

specialization_data <- read_csv("SPECIALIZATION_DATA.csv")
# remove variables: Space, Initial Admit Number, Inital Reject Number (irrevelevant to analysis) & change column names 
spec_data <- specialization_data[,-c(4,5,6,7)]
colnames(spec_data) <- c("Year","Option","Spec","max_grade","min_grade")

# remove rows that have no specialization or have no min grade
spec_data <- filter(spec_data, !is.na(spec_data$Spec) & !is.na(spec_data$min_grade))
spec_data <- filter(spec_data, !spec_data$min_grade== 0)

# Turn Max and Min Grade from chr to dbl
spec_data$min_grade <- as.numeric(spec_data$min_grade)
spec_data$max_grade <- as.numeric(spec_data$max_grade)

# Ensure max grade > min grade to make sure there is no data entry errors 
#filter(spec_data, (spec_data$min_grade > spec_data$max_grade)) # found biotech in 2020 min was 89.25 while max for 73.96 so I switch them <- Test  
spec_data$min_grade[(spec_data$Year == 2020) & (spec_data$Spec == "Honours (1136): Biotechnology")] <- 73.96
spec_data$max_grade[(spec_data$Year == 2020) & (spec_data$Spec == "Honours (1136): Biotechnology")] <- 89.25
# filter(spec_data, (spec_data$Year == 2020) & (spec_data$Spec == "Honours (1136): Biotechnology")) # <- Test

# Found that CPSC (Excluding CPSC Major Domestic Student) & CPSC (Excluding Domestic Student) are the same thing but different names --> I also change the names to make more sense
spec_data$Option[(spec_data$Option == 'CPSC (Excluding CPSC Major Domestic Student)')] <- 'CPSC (Excluding Domestic Student)'

spec_data$Option[(spec_data$Option == 'CPSC (Excluding Domestic Student)')]       <- 'CPSC (International)'
spec_data$Option[(spec_data$Option == 'CPSC (Excluding International Students)')] <- 'CPSC (Domestic)'

# Found that 'Honours (1093): Biophysics (BIOP)' option switches from Honours to Combined Honours in 2020, I switch it to Honours for consistency 
spec_data$Option[((spec_data$Year == 2020) | (spec_data$Year == 2021)) & (spec_data$Spec == 'Honours (1093): Biophysics (BIOP)')] <- 'Honours'
#filter(spec_data, spec_data$Spec == 'Honours (1093): Biophysics (BIOP)') # <- Test  

# Same with The Options: Combined Majors & Combined Major
spec_data$Option[(spec_data$Option == 'Combined Majors')] <- 'Combined Major'

# I also add a dummy variable to identify which years had COVID's affect (2020 and 2021)
spec_data <- spec_data %>%
  mutate(COVID = case_when(
    ((Year == 2020 | Year == 2021)) ~ 1,
     TRUE ~ 0)) %>%
      as_factor()

print(spec_data)
```

In 2018, I found CPSC splits from general seats to having seperate seats and therefore admission requirements for domestic and international students. Because of this in a lot of my analysis, I'll be combining the two together as an aggregate in order to have a consistent measure of CPSC specialization requirements. The following code block is utilized throughout all the parts to make use of this aggregate.

```{r Aggregating Comp Sci}
# this code box is nesessary
CPSC_Data = filter(spec_data, spec_data$Option == 'CPSC' | spec_data$Option == 'CPSC (Domestic)' | spec_data$Option == 'CPSC (International)')
CPSC_Data

CPSC_Min <- CPSC_Data %>%
            group_by(Year) %>%
            summarize(
              Option = 'CPSC', 
              Spec = 'CPSC',
              max_grade = mean(max_grade),
              min_grade = mean(min_grade),
            ) 

CPSC_Min <- CPSC_Min %>%
  mutate(COVID = case_when(
    ((Year == 2020 | Year == 2021)) ~ 1,
     TRUE ~ 0)) %>%
      as_factor()

#Use Interpolation to estimate 2018 min grade as there is no data on domestic min grade 
CPSC_2018_mean <- mean(c((CPSC_Min$min_grade[CPSC_Min$Year == 2017]),(CPSC_Min$min_grade[CPSC_Min$Year == 2019])))

CPSC_Min$min_grade[CPSC_Min$Year == 2018] <- CPSC_2018_mean
CPSC_Min$Option <- 'Aggregated CPSC'
CPSC_Min$Spec <- 'Aggregated CPSC'


CPSC_Min
spec_data <- rbind(spec_data, CPSC_Min)

spec_data
```


# Part 2: Looking at individual Options (Average of Everything, CPSC, HONOURS, Integrated Science (specific major known for being premed))

The following code blocks can be used for initial exploration on how the avg_min for certain "options" of specializations have changed from 2015 - 2021. These options are from UBC's data itself to categorize specializations in terms of honours, major, CPSC, etc. 

```{r Summarize - All Majors}
Yearly_Min <- spec_data %>%
            group_by(Year) %>%
            summarize(
                avg_min = mean(min_grade),
                avg_max = mean(max_grade)
            )

Yearly_Min
```

```{r Summarize - Honours}
Honours_data <- filter(spec_data, spec_data$Option == 'Honours')
Honours_min <- Honours_data %>% 
  group_by(Year) %>% 
  summarize(
    avg_min = mean(min_grade),
    avg_max = mean(max_grade)
  )

Honours_min
```

```{r Summarize - Integrated Science}
integrated_data <- filter(spec_data, spec_data$Spec== 'Major (1682): Integrated Sciences (INSC)')
integrated_data
integrated_min <- integrated_data %>% 
  group_by(Year) %>% 
  summarize(
    avg_min = mean(min_grade),
    avg_max = mean(max_grade)
  )

integrated_min
```

# Part 3: Looking at Relationships & Visual Plots

I first begin by creating a plot that showcases the relationship between different 'Options' (that is defined by UBC) and how their minimum requirements change throughout the years 

I'll first work the data in order to let it work in my visualization

```{r Pre-emptive Cleaning up of data}
# CPSC splits into Domestic & International in 2018, so for the sake of consistency I make a new Option that combines them 
spec_options_data <- spec_data 

spec_options_data <- arrange(spec_options_data, Year, Option)
spec_options_data <- filter(spec_options_data, !(spec_options_data$Option == 'CPSC (International)' | spec_options_data$Option == 'CPSC (Domestic)' |  spec_options_data$Option == 'CPSC')) # Remove CPSC 


# I combine COGS and STATS into it's own option to avoid clutter, also noting that pre-2017, these specializations had their option as 'Major" so I set them appropiately in the data set
spec_options_data$Option[(spec_options_data$Spec == 'Major Cognitive Systems(1225): Cognition and Brain')] <- 'COGS'
spec_options_data$Option[(spec_options_data$Spec == 'Major (0562): Statistics (STAT)')] <- 'STAT'

spec_options_data$Option[(spec_options_data$Option == 'COGS')] <- 'COGS/STATS'
spec_options_data$Option[(spec_options_data$Option == 'STAT')] <- 'COGS/STATS'

```

Then I make the visualization, noting that the CPSC category after 2018 is an average of the domestic/international requirements

```{r Plot of UBC Options Requirements}
options_data <- spec_options_data %>%
  group_by(Option, Year) %>%
  summarize(
    avg_min = mean(min_grade),
    avg_max = mean(max_grade),
    avg_sd = sd(min_grade))
  
  options_data
  
  plot <- ggplot(data = options_data,
  aes(
  x = Year,
  y = avg_min,
  color = Option,
  fill = Option
  ))
  
  
  plot <- plot + labs(
  x = "Year",
  y = "Grade Requirement",
  color = "Specialization Option",
  fill = "Specialization Option",
  title = "Grade Requirement over Time"
  )
  
  plot <- plot + scale_color_brewer(palette ="Dark2")
    
  plot1 <- plot + geom_line() + geom_vline(xintercept = 2020, linetype = "dotted") + annotate(geom="label", x=2020, y=50, label="COVID Begins", color="black")
  plot1
```
This plot demonstrates the trend in which many of the Specialization Options have increasing grade requirements as time goes on. It is interesting to note that from 2015 to 2021, the "Combined Major" option has no long term change. As these options have little meaning due to their inconsistencies in what field the specalizations are in (except CPSC & COGS/STATS), the competetiveness of these options are extremely varied because of it. We can see that as we look at the standard deviation for each option, options such as "Major" and "Combined Major" are very volatile and therefore the mean quite unreliable in representation. Because of this, I will opt to make my own categories to categorize specializations. 

# Part 3b: Looking at Grade Requirements for Custom Categories 

In the following plot, instead of using UBC's 'options' to aggregate specializations, I instead use my own categories in which I believe can provide a better representation of the data. The categories I choose are: Life Sciences, Numerical Sciences, Computer Science (it's own specialization), Specializations considered competetive in 2021 (have a requirement of 75+) and an aggregate of all the specializations


```{r Custom Categories & Requirements}
#This function creates a subset of spec_data based on the list of specializations given 
list_spec <- function(speclist) {
  listofspec <- c() #gives me an empty subset 
  
  for (val in speclist) {
    specific_subset <- subset(spec_data, Spec == val)
    listofspec <- rbind(listofspec, specific_subset)
  }
  return(listofspec)
}

#This lists is determined by how UBC defines programs on their website (https://you.ubc.ca/programs/#mode=by-topic&viewMode=list) 
lifesciences <- c('Major (3502): Behavioural Neuroscience (PSYC)', 'Major (0244): Biochemistry (BIOC)', 'Major (3095): Biology (BIOL)', 'Major (1682): Integrated Sciences (INSC)', 'Major (1153): Microbiology and Immunology (MBIM)', 'Major (0311): Pharmacology (PCTH)', 'Honours (1136): Biotechnology', 'Honours (3221): Cellular, Anatomical and Physiological Sciences (CAPS)', 'Combined Honours (3508): Biochemistry and Forensic Science', 'Honours (1093): Biophysics (BIOP)')

#This list is what I consider numerical sciences (Note: I use the average of International and Domestic CPSC after the specialization splits in 2018)
numericalsciences <- c('Major (0562): Statistics (STAT)', 'Major Cognitive Systems(1225): Cognition and Brain', 'Major Cognitive Systems(1226): Computational Intelligence and Design', 'Major (0456): Mathematics (MATH)', 'Major (0524) Physics (PHYS)', 'Honours (0166): Mathematics (MATH)', 'Honours (0344): Physics (PHYS)', 'Combined Honours (0009): Physics and Astronomy (PHYS, ASTR)', 'Combined Honours (0014): Physics and Mathematics (PHYS, MATH)') 

#Assigning Categories to each subset 
categorized_data_lifesciences <- list_spec(lifesciences) %>%
  mutate(Category = 'Life Sciences') %>% 
        as_factor()

categorized_data_numericalsciences <- list_spec(numericalsciences) %>%
  mutate(Category = 'Numerical Sciences') %>% 
        as_factor()

categorized_data_compsci <- list_spec(c("Aggregated CPSC")) %>%
  mutate(Category = 'Computer Science') %>% 
        as_factor()

#This a list of specializations that have a requirement >75 in 2021 
spec_data_comp_2021 <- spec_data %>%
    filter((Year == 2021) & (min_grade > 75) & !Option == 'CPSC (Domestic)' & !Option == "CPSC (International)")

list_spec_comp <- c(t(spec_data_comp_2021$Spec))

categorized_data_competetive<- list_spec(list_spec_comp) %>%
      mutate(Category = 'Comp Specializations (Including CPSC)') %>% 
      as_factor()

categorized_data_everything <- spec_data %>%
    mutate(Category = 'All Specializations') %>% 
      as_factor()

categorized_data <- rbind(categorized_data_lifesciences, categorized_data_numericalsciences, categorized_data_compsci, categorized_data_competetive, categorized_data_everything)

categorized_data

#Plot 
categorized_data_plot <- categorized_data %>%
  group_by(Category, Year) %>%
  summarize(
    avg_min = mean(min_grade),
    avg_max = mean(max_grade))
  
  categorized_data_plot
  
  plot <- ggplot(data = categorized_data_plot,
  aes(
  x = Year,
  y = avg_min,
  color = Category,
  fill = Category,
  label = Category
  ))
  
  
  plot <- plot + labs(
  x = "Year",
  y = "Grade Requirement",
  color = "Degree Category",
  fill = "Degree Category",
  title = "Grade Requirement over Time"
  )
  
  plot <- plot + scale_color_brewer(palette ="Dark2")  
  categorized_plot <- plot + geom_line() + geom_vline(xintercept = 2020, linetype = "dotted") + annotate(geom="label", x=2020, y=62.5, label="COVID Begins", color="black")
  categorized_plot 
```
The plot definetely shows that specializations have requirements rising throughout the years, in particular in 2020 and 2021 which are referred as the COVID years. Something to note though is that scale of the plot is only in 5% intervals in which most of the changes are within. These are not crazy changes and therefore cannot be conclusive on their own. In the following part, I'll use statistics to determine how probable this situation would've been following the trends in 2015-2019. 

In table form, this data looks like this: 

```{r}
summarized_data_lifesciences <- categorized_data_lifesciences %>%
  group_by(Year) %>%
  summarize(
    life_sciences_min = mean(min_grade))

summarized_data_numericalsciences <- categorized_data_numericalsciences %>%
  group_by(Year) %>%
  summarize(
    numerical_sciences_min = mean(min_grade))

summarized_data_comp <- categorized_data_competetive%>%
  group_by(Year) %>%
  summarize(
    competetive_specializations_min = mean(min_grade))

summarized_data_cpsc <- categorized_data_compsci %>%
  group_by(Year) %>%
  summarize(
    comp_sci_min = mean(min_grade))

summarized_data_allspecializations <- categorized_data_everything %>%
  group_by(Year) %>%
  summarize(
    all_specializations_min = mean(min_grade))

summarized_data <- merge(summarized_data_lifesciences, summarized_data_numericalsciences, by = "Year")
summarized_data <- merge(summarized_data, summarized_data_comp, by = "Year")
summarized_data <- merge(summarized_data, summarized_data_cpsc, by = "Year")
summarized_data <- merge(summarized_data, summarized_data_allspecializations, by = "Year")

summarized_data
```

# Part 4: Actual Statistics to see if COVID affected requirements
Using the data made in Part 3b (categorized_data) which is the data on requirement averages of the various categories I made, I will go on to use linear regression in order to extrapolate and forecast what 2020 & 2021 would be with the trends present in non-covid years and use that to compare with the actual 2020 & 2021 averages that were affected by COVID. Due to the inconsistency in 2015-2016 in relative to the actual mean, I'll be not including these in the regression line. 

```{r Regression Plots of Categorized Data}
#This function makes a regression plot of a categorized dataset
make_regression_plot <- function(categorized_dataset) {
categorized_dataset_plot_data <- categorized_dataset %>%
  group_by(Category, Year) %>%
  summarize(
    avg_min = mean(min_grade),
    avg_max = mean(max_grade))

plot <- ggplot(data = categorized_dataset_plot_data,
aes(
  x = Year,
  y = avg_min,
  color = Category,
  fill = Category
  ))
  
  
plot <- plot + labs(
  x = "Year",
  y = "Grade Requirement",
  color = "Degree Category",
  fill = "Degree Category",
  title = "Grade Requirement over Time"
  )
  

filtered_data <- filter(categorized_dataset_plot_data, (!categorized_dataset_plot_data$Year == 2020 & !categorized_dataset_plot_data$Year == 2021 & !categorized_dataset_plot_data$Year == 2015 & !categorized_dataset_plot_data$Year == 2016)) 
# filtering here makes the regression line not include 2020 & 2021 and 2015 & 2016

reg <- lm(avg_min~Year, filtered_data)
coeff <- coefficients(reg)

intercept <- coeff[1]
slope <- coeff[2]

categorized_plot <- plot + geom_point() + geom_abline(slope = slope, intercept = intercept) + geom_vline(xintercept = 2020, linetype = "dotted") + annotate(geom="label", x=2020, y=62.5, label="COVID Begins", color="black") + ylim(60,85)
return (categorized_plot)
}

categorized_datasets <- list(categorized_data_lifesciences, categorized_data_numericalsciences, categorized_data_competetive, categorized_data_compsci, categorized_data_everything)

regression_plots <- lapply(categorized_datasets, make_regression_plot)
regression_plots
```

After seeing the regression plots, it can be seen that the 2020 & 2021 requirements do seem to generally fit in with the projected means. I can further dive into this using a t-test involving all the categories except comp sci (because there is not enough observations) by analyzing if the actual mean in 2021 is probable with 95% confidence to the projected 2021 mean that is derived from the regression analysis from the previous part. 

Null hypothesis: The projected requirement mean of 2021 of the regression line equals the actual requirement mean
Alternative hypothesis: The projected requirement mean of 2021 of the regression line does not equal the actual requirement mean

```{r T-Tests of Categorized Data}

categorized_datasets <- list(categorized_data_lifesciences, categorized_data_numericalsciences, categorized_data_competetive, categorized_data_everything)

#This function takes in a dataset and does a t-test to evaluate the acutal 2021 requirement mean in comparison to the projected mean from the regression line that extrapolates from data in 2015 - 2019
t_test_function <- function(categorized_dataset) {

categorized_dataset_plot_data <- categorized_dataset %>%
  group_by(Category, Year) %>%
  summarize(
    avg_min = mean(min_grade),
    sd_min = sd(min_grade),
    avg_max = mean(max_grade))
  
filtered_data <- filter(categorized_dataset_plot_data, (!categorized_dataset_plot_data$Year == 2020 & !categorized_dataset_plot_data$Year == 2021 & !categorized_dataset_plot_data$Year == 2015 & !categorized_dataset_plot_data$Year == 2016)) 
# filtering here makes the regression line not include 2020 & 2021 and 2015 & 2016

reg <- lm(avg_min~Year, filtered_data)
coeff <- coefficients(reg)

intercept <- coeff[1]
slope <- coeff[2]

projected2021 <- intercept + slope*2021

# Hypothesis Test done by hand

# actual2021 <- categorized_dataset_plot_data$`avg_min`[categorized_dataset_plot_data$Year == 2021]
# actualSD2021 <- categorized_dataset_plot_data$`sd_min`[categorized_dataset_plot_data$Year == 2021]

# test <- pnorm(projected2021, actual2021, actualSD2021)
# test


# Confidence Intervals done by hand
# number_of_rows <-  categorized_dataset %>%
#   group_by(Spec) %>%
#   summarize() %>% 
#   nrow()
#   
#   
# standarderror <- actualSD2021 / sqrt(number_of_rows)
# criticalvalue <- qt(((1-0.95)/2), 9)
# 
# confidenceintervals <-- c((actual2021-(criticalvalue*-1)*standarderror)*-1, (actual2021+(criticalvalue*-1)*standarderror)*-1)
# 
# confidenceintervals

filtered <- filter(categorized_dataset, categorized_dataset$Year == 2021)$min_grade
filtered


t <- t.test(
      filtered,
       alternative = "two.sided",
       mu = projected2021,
       conf.level = 0.95)

return (t)
}

t_tests <- lapply(categorized_datasets, t_test_function)
t_tests

#In order the T-tests are: lifesciences, numericalsciences, competetive_specializations, all_specializations
```

The first t-test being life sciences showcases accepting the null hypothesis while the others reject it. This can display an influence of COVID on the specialization requirements, and is something also seen in the regression plots. 

# Part 4b: Pair T-Test of Pre-Covid & Post-Covid
Another signifiance test I can do is to compare the requirement means of pre-covid years and post-covid years. To be representative, I'll look at a symmetric window between non-covid years and covid years (that being 2017 & 2018 vs 2020 & 2021) in order to have a representative comparison of the trend. 

```{r Pair T-Tests of Pre-Post COVID}
# Do a pair T test of Pre Covid and Post COVID using the COVID dummy variable to compare if the min requirements are equal

categorized_datasets <- list(categorized_data_lifesciences, categorized_data_numericalsciences, categorized_data_competetive, categorized_data_everything)

#This function takes in a dataset and does a t-test to evaluate the acutal 2021 requirement mean in comparison to the projected mean from the regression line that extrapolates from data in 2015 - 2019
t_test_function <- function(categorized_dataset) {
  
  filtered_data <- filter(categorized_dataset, (!categorized_dataset$Year == 2015 & !categorized_dataset$Year == 2016 & !categorized_dataset$Year == 2019)) 
  
  t <- t.test(
    x = filter(filtered_data, COVID == 0)$min_grade,
    y = filter(filtered_data, COVID == 1)$min_grade,
       
    alternative = "two.sided",
    mu = 0,
    conf.level = 0.95)

  return (t)
}

t_tests <- lapply(categorized_datasets, t_test_function)
t_tests

#In order the T-tests are: lifesciences, numericalsciences, competetive_specializations, all_specializations
```

Again what we can see here is life sciences accepting the null hypothesis while the rest reject. This showcases COVID as a factor in the rise of specialization requirements

# Part 5: Alternative View: Looking at how "Competetive" UBC Specializations are getting by Proportion 

The following plot will demonstrate the proportion of specializations that have requirements above a certain cutoff. By viewing how this proportion changes throughout the years, it can be seen how competeteive specializations are getting.  

```{r Proportion of Specializations > Cutoff Plot}
#This function takes in a cutoff and provides a table of how many specializations per year have a min grade above that cutoff 
proportion_spec_with_cutoff <- function(cutoff) {
comp_spec_data <- spec_data %>% 
    mutate(Competetive = case_when( 
        (min_grade > cutoff) ~ "Competetive",
        TRUE ~ "Uncompetetive")) %>% 
        as_factor()
 
    competetive_spec_data <- comp_spec_data
    count_table <- competetive_spec_data %>%
    group_by(Year, Competetive) %>%
    summarize( 
    'Count' = n(),
    Cutoff = cutoff)
    count_table$Cutoff <- as.character(count_table$Cutoff)
    
    count_table
    
    aggregated_count_table <- c()
    
# Calculates total amount of specializations per year to have proportion per year instead of proportion of total amount of specializations due to UBC adding/removing specializations
    for(i in 2015:2021) {
    count_table_specific_year <- count_table %>%
      ungroup %>%
      filter(Year == i)
    
    count_table_specific_year <- count_table_specific_year %>% 
      mutate(total_year_count = sum(count_table_specific_year$Count))
    
    aggregated_count_table <- rbind(aggregated_count_table, count_table_specific_year)
    }
  
    competetive_spec_data <- filter(aggregated_count_table, aggregated_count_table$Competetive == "Competetive")
    proportion_table <- competetive_spec_data %>%
    group_by(Year) %>%
    summarize( 
    'Proportion' = Count / total_year_count,
    Cutoff = cutoff)
    proportion_table$Cutoff <- as.character(proportion_table$Cutoff)
    
return (proportion_table)
}

proportion_data <- rbind(proportion_spec_with_cutoff(65), proportion_spec_with_cutoff(70), proportion_spec_with_cutoff(75), proportion_spec_with_cutoff(80), proportion_spec_with_cutoff(85))

proportion_data

proportion_plot <- ggplot(proportion_data,
                       aes(
                         x = Year,
                         y = Proportion*100,
                         color = Cutoff,
                         fill = Cutoff
                       ))

proportion_plot <- proportion_plot + labs(
  x = "Year",
  y = "% of Specializations",
  color = "Cutoff",
  fill = "Cutoff",
  title = "Proportion of Specializations Over a Cutoff"
  )
    
  proportion_plot <- proportion_plot + geom_col(position = "dodge") 
  
  proportion_plot
```

In 2020 and 2021, there is an increase in the proportions drastically in the cutoffs of 75 & 80 implying that many specializations that were previously not as competetive became such during COVID. 

This can be further viasualized by looking at difference in proportion in the years vs 2015 rather than total proportion

```{r Difference in Proportion from 2015}
#This function takes in a cutoff and provides a table of how many specializations per year have this requirement more there is than in 2015 
proportion_diff_with_cutoff <- function(cutoff) {
 proportion_table <- proportion_spec_with_cutoff(cutoff)
 diff_table <- proportion_table %>%
  mutate(Difference = Proportion - proportion_table$Proportion[Year == 2015]) %>%
  filter(!proportion_table$Year == 2015)
  
return (diff_table)
}

diff_data <- rbind(proportion_diff_with_cutoff(65), proportion_diff_with_cutoff(70), proportion_diff_with_cutoff(75), proportion_diff_with_cutoff(80), proportion_diff_with_cutoff(85))

diff_data

diff_plot <- ggplot(diff_data,
                       aes(
                         x = Year,
                         y = Difference*100,
                         color = Cutoff,
                         fill = Cutoff
                       ))

diff_plot <- diff_plot + labs(
  x = "Year",
  y = "Difference in Proportion from 2015",
  color = "Cutoff",
  fill = "Cutoff",
  title = "Difference in Proportion of Specializations Over a Cutoff from 2015"
  )
    
  diff_plot <- diff_plot + geom_col(position = "dodge")
  
  diff_plot
```

The increase in specializations with a requirement > 75 & 80 can be viewed more representatively with this visualization and it's implications of COVID making specializations more competetive. 

# Part 5b - Looking at Competetetive Specializations but with Consistent Specializations
Though the plots in part 5 demonstrate a change in proportion, it does not account for the fact that UBC adds/removes majors throughout the years. This could cause issues such as UBC adding more competetive specializations in 2020 & 2021 that can explain the increase in proportion in those years. Therefore, in the following analysis, I will only look at specializations that are present throughout 2015 - 2021 consistently in the following plots. Due to CPSC being split in 2018, I'll also be using the average of domestic/international

```{r Creating Consistent Data}
spec_data_cpsc_consistent <- filter(spec_data, (!spec_data$Option == 'CPSC (International)' & !spec_data$Option == 'CPSC (Domestic)' & !spec_data$Option == 'CPSC'))
spec_data_cpsc_consistent <- rbind(spec_data_cpsc_consistent)
spec_data_consistent <- spec_data_cpsc_consistent %>%
  group_by(Spec) %>%
  filter(all(2015:2021 %in% Year))

  total_num_specializations_consistent <- spec_data_consistent %>%
    group_by(Spec) %>%
    summarize() %>%
    nrow()

#This is a test to verify that the specializations are consistent throughout the years
proportion_spec_data <- spec_data_consistent %>%
    group_by(Year) %>%
    summarise(Proportion = n())
    
proportion_spec_data

spec_data_consistent

```

The following code blocks will be repeating the previous visualizations in Part 5 but instead of looking at all the specializations, I'll only be looking at specializations present throughout 2015-2021 consistently. 

```{r Specialization Proportion Above Cutoff (Consistent)}
proportion_spec_with_cutoff_consistent <- function(cutoff) {
    comp_spec_data_consistent <- spec_data_consistent %>% 
    mutate(Competetive = case_when( 
        (min_grade > cutoff) ~ "Competetive",
        TRUE ~ "Uncompetetive")) %>% 
        as_factor()
 
    competetive_spec_data_consistent <- comp_spec_data_consistent
    count_table_consistent <- competetive_spec_data_consistent %>%
    group_by(Year, Competetive) %>%
    summarize( 
    'Count' = n(),
    Cutoff = cutoff)
    count_table_consistent$Cutoff <- as.character(count_table_consistent$Cutoff)
    
    count_table_consistent
    
    aggregated_count_table_consistent <- c()
    
# Calculates total amount of specializations per year to have proportion per year instead of proportion of total amount of specializations due to UBC adding/removing specializations
    for(i in 2015:2021) {
    count_table_specific_year_consistent <- count_table_consistent %>%
      ungroup %>%
      filter(Year == i)
    
    count_table_specific_year_consistent <- count_table_specific_year_consistent %>% 
      mutate(total_year_count = sum(count_table_specific_year_consistent$Count))
    
    aggregated_count_table_consistent <- rbind(aggregated_count_table_consistent, count_table_specific_year_consistent)
    }
  
    competetive_spec_data_consistent <- filter(aggregated_count_table_consistent, aggregated_count_table_consistent$Competetive == "Competetive")
    proportion_table_consistent <- competetive_spec_data_consistent %>%
    group_by(Year) %>%
    summarize( 
    'Proportion' = Count / total_year_count,
    Cutoff = cutoff)
    proportion_table_consistent$Cutoff <- as.character(proportion_table_consistent$Cutoff)
    
return (proportion_table_consistent)
}

proportion_data_consistent <- rbind(proportion_spec_with_cutoff_consistent(65), proportion_spec_with_cutoff_consistent(70), proportion_spec_with_cutoff_consistent(75), proportion_spec_with_cutoff_consistent(80), proportion_spec_with_cutoff_consistent(85))
# to test if this is actually consistent, put cutoff at 0 and add it onto data
proportion_data_consistent

proportion_plot_consistent <- ggplot(proportion_data_consistent, 
                       aes(
                         x = Year,
                         y = Proportion*100,
                         color = Cutoff,
                         fill = Cutoff
                       ))

proportion_plot_consistent <- proportion_plot_consistent + labs(
  x = "Year",
  y = "% of Specializations",
  color = "Cutoff",
  fill = "Cutoff",
  title = "Proportion of Specializations Over a Cutoff (Consistent)"
  )
    
  proportion_plot_consistent <- proportion_plot_consistent + geom_col(position = "dodge")
  
  proportion_plot_consistent
```

And for difference in Proportion again:

```{r Specialization Proportion Difference from 2015 (Consistent)}
proportion_diff_with_cutoff_consistent <- function(cutoff) {
 proportion_table_consistent <- proportion_spec_with_cutoff_consistent(cutoff)
 diff_table <- proportion_table_consistent %>%
  mutate(Difference = Proportion - proportion_table_consistent$Proportion[Year == 2015]) %>%
  filter(!proportion_table_consistent$Year == 2015)
  
return (diff_table)
}

diff_data_consistent <- rbind(proportion_diff_with_cutoff_consistent(65), proportion_diff_with_cutoff_consistent(70), proportion_diff_with_cutoff_consistent(75), proportion_diff_with_cutoff_consistent(80), proportion_diff_with_cutoff_consistent(85))

diff_data_consistent

diff_plot_consistent <- ggplot(diff_data_consistent,
                       aes(
                         x = Year,
                         y = Difference*100,
                         color = Cutoff,
                         fill = Cutoff
                       ))

diff_plot_consistent <- diff_plot_consistent + labs(
  x = "Year",
  y = "Difference in Proportion from 2015",
  color = "Cutoff",
  fill = "Cutoff",
  title = "Difference in Proportion of Specializations Over a Cutoff from 2015 (Consistent)"
  )
    
  diff_plot_consistent <- diff_plot_consistent + geom_col(position = "dodge")
  
  diff_plot_consistent
```

I can see the difference this makes by comparing the plots side by side

```{r Graphs side by side}
proportion_plot + ylim(0,100)
proportion_plot_consistent + ylim(0,100)
diff_plot + ylim(-50,50)
diff_plot_consistent + ylim(-50,50)
```

It can be seen that the difference between these two plots highlight that the proportion of specializations with a requirement > 75 is much less volatile until 2019 in the consistent plot which is interesting as 2019 had no COVID affect. We can also see that the cutoff above 85 has drastically less change in the consistent plot. In order to understand more on these changes in cutoffs in this year, the following code blocks provide a function that allows us to see which specific specializations crossed this cutoff from a lower year to an upper year. 

```{r Making Function to Find Specializations with Dramatic Change} 
#This function takes in 2 years and returns a table of specializations that in the lower year, had a requirement lower than the cutoff but in the upperyear had a requirement above the cutoff 
specialization_requirement_change <- function(loweryear, upperyear, cutoff) {

spec_data_consistent_part4 <- spec_data %>%
  group_by(Spec) %>%
  filter(all(loweryear:upperyear %in% Year))

spec_data_consistent_part4

summarise_spec_spec_data <- spec_data_consistent_part4 %>%
  group_by(Spec) %>%
  summarise()

summarise_spec_spec_data

list_spec_data <- c(t(summarise_spec_spec_data$Spec))

list_spec_data <- list_spec_data
list_spec_data

empty_list_spec <- c()
for (variable in list_spec_data) {
  lowerrequirement <- spec_data$min_grade[(spec_data$Year == loweryear) & (spec_data$Spec == variable)]
  upperrequirement <- spec_data$min_grade[(spec_data$Year == upperyear) & (spec_data$Spec == variable)]
  requirement <- (lowerrequirement < cutoff & cutoff <= upperrequirement)
  
  if(requirement) {
    specific_subset <- subset(spec_data, ((spec_data$Spec == variable) & ((spec_data$Year == loweryear) | (spec_data$Year == upperyear))))
    empty_list_spec <- rbind(empty_list_spec, specific_subset)
  }
}

return (empty_list_spec)
}
```

In particular, a relatively spike can be found between 2018 and 2019 in that there is a large Proportion difference for majors with a requriement > 75. By using this tool I can isolate 5 majors that showcase this change and using the tool in the reverse order, can find 2 specializations that show the reverse

```{r}
#Needs year 2015 - 2021 
specialization_requirement_change(2018, 2019, 75)
specialization_requirement_change(2019, 2018, 75)
```
While this does not implicate any correlation towards COVID, it is interesting to see which specializations are drastically changing through the years. 

#Part 5c: Interactive Code! 

This function takes in 2 years and returns a table of specializations that in the first year, had a requirement lower than the cutoff but in the second year had a requirement above the cutoff 

```{r}
specialization_requirement_change(2019, 2021, 75)
```



